---
title: "Quantium VI Task 2"
author: "Rolly Mougoue"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(linewidth=80)
```

```{r setup2, include=FALSE}
library(knitr)
hook_output = knit_hooks$get("output")
knit_hooks$set(output = function(x, options)
{
# this hook is used only when the linewidth option is not NULL
if (!is.null(n <- options$linewidth))
{
x = knitr:::split_lines(x)
# any lines wider than n should be wrapped
if (any(nchar(x) > n))
x = strwrap(x, width = n)
x = paste(x, collapse = "\n")
}
hook_output(x, options)
})
```

# Task 2 : Retail Strategy and Analytics

## Load required libraries and datasets

We'll start with setting up our working environment by installing and loading all the necessary packages (Tidyverse, Lubridate, dplyr, pad...) and setting the working directory for easy upload of our datasets

```{r Load libraries, include = FALSE}
library(tidyverse)
library(data.table)
library(ggplot2)
library(ggmosaic)
library(readr)
library(tidyr)


## Setting the work directory and uploading our data sets


setwd(dir = "D:/Google Data Analytics/The Forage Projects/Quantium/Datasets")


Data <- read_csv("QVI_data.csv")
```



## Set themes for plots
```{r Setting Plot Themes}
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
Data <- data.table(Data)
```

```{r Viewing Dataset}
show(Data)
```

## Select control stores

The client has selected store numbers 77, 86 and 88 as trial stores and want control stores to be established stores that are operational for the entire observation period.We would want to match trial stores to control stores that are similar to the trial
store prior to the trial period of Feb 2019 in terms of :
** Monthly overall sales revenue
** Monthly number of customers
** Monthly number of transactions per customer
Let's first create the metrics of interest and filter to stores that are present throughout the pre-trial period.

```{r Select control stores}
#### Calculate these measures over time for each store 
## Add a new month ID column

Data[, YEARMONTH := year(DATE)*100 + month(DATE)]

#### Next, we define the measure calculations

measureOverTime <- Data[, .(totSales = sum(TOT_SALES) ,
 nCustomers = uniqueN(LYLTY_CARD_NBR),
 nTxnPerCust = uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR),
 nChipsPerTxn = sum(PROD_QTY)/uniqueN(TXN_ID),
 avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY)),
 by = c("STORE_NBR", "YEARMONTH")][order(STORE_NBR, YEARMONTH)]
```

```{r}
measureOverTime
```


```{r }
#### Filter to the pre-trial period and stores with full observation periods
storesWithFullObs <- unique(measureOverTime[, .N, STORE_NBR][N == 12, STORE_NBR])
storesWithFullObs
```

```{r }
preTrialMeasures <- measureOverTime[YEARMONTH < 201902 & STORE_NBR %in% storesWithFullObs, ]
preTrialMeasures
```

Now we need to work out a way of ranking how similar each potential control store is to the trial store. We can calculate how correlated the performance of each store is to the trial store.
Let's write a function for this so that we don't have to calculate this for each trial store and control store pair.

```{r Create function to calculate correlation}
#### Let's define inputTable as a metric table with potential comparison stores, metricCol as the store metric used to calculate correlation on, and storeComparisonas the store number of the trial store.

calculateCorrelation <- function(inputTable, metricCol, storeComparison) {
  
  calcCorrTable = data.table(Store1 = numeric(), Store2 = numeric(), corr_measure = numeric())
  
  storeNumbers <- unique(inputTable[, STORE_NBR])
  
  for (i in storeNumbers) {
    calculatedMeasure = data.table("Store1" = storeComparison, "Store2" = i,
                                   "corr_measure" = cor( inputTable[STORE_NBR == storeComparison,
                                                                   eval(metricCol)], inputTable[STORE_NBR == i,
                                                                                              eval(metricCol)]))
    
    calcCorrTable <- rbind(calcCorrTable, calculatedMeasure)
  }
  
return(calcCorrTable)
}

```

Apart from correlation, we can also calculate a standardised metric based on the absolute difference between the trial store's performance and each control store's performance.
Let's write a function for this.

```{r Create function to calculate magnitude distance}
calculateMagnitudeDistance <- function(inputTable, metricCol, storeComparison){
  calcDistTable = data.table(Store1 = numeric(), Store2 = numeric(), YEARMONTH =
                               numeric(), measure = numeric())
  
  storeNumbers <- unique(inputTable[, STORE_NBR])
  
  for (i in storeNumbers) {
    calculatedMeasure = data.table("Store1" = storeComparison, "Store2" = i, 
                                   "YEARMONTH" = inputTable[STORE_NBR == storeComparison, YEARMONTH], 
                                   "measure" = abs(inputTable[STORE_NBR == storeComparison, eval(metricCol)]
                                                   - inputTable[STORE_NBR == i,eval(metricCol)])
                                   )
    calcDistTable <- rbind(calcDistTable, calculatedMeasure)
  }

#### Standardise the magnitude distance so that the measure ranges from 0 to 1
minMaxDist <- calcDistTable[, .(minDist = min(measure), maxDist = max(measure)), by = c("Store1","YEARMONTH")]
  
  distTable <- merge(calcDistTable, minMaxDist, by = c("Store1", "YEARMONTH"))
  distTable[, magnitudeMeasure := 1 - (measure - minDist)/(maxDist - minDist)]
  
  finalDistTable <- distTable[, .(mag_measure = mean(magnitudeMeasure)), by = .(Store1, Store2)]
  return(finalDistTable)
}

    
```

Now let's use the functions to find the control stores! We'll select control stores based on how similar monthly total sales in dollar amounts and monthly number of customers are to the trial stores. So we will need to use our functions to get four scores, two for each of total sales and total customers.

```{r Use functions to calculate metrics}
#### Calculate correlations against store 77 using total sales and number of customers.
trial_store <- 77

corr_nSales <- calculateCorrelation(preTrialMeasures, quote(totSales), trial_store)
corr_nSales[order(-corr_measure)]

corr_nCustomers <- calculateCorrelation(preTrialMeasures, quote(nCustomers), trial_store)
corr_nCustomers[order(-corr_measure)]


#### Then, use the functions for calculating magnitude.
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales), trial_store)
magnitude_nSales[order(-mag_measure)]

magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures, quote(nCustomers), trial_store)
magnitude_nCustomers[order(-mag_measure)]
```

We'll need to combine the all the scores calculated using our function to create a composite score to rank on. 
Let's take a simple average of the correlation and magnitude scores for each driver. Note that if we consider it more important for the trend of the drivers to be similar, we can increase the weight of the correlation score (a simple average gives a weight of 0.5 to the corr_weight) or if we consider the absolute size of the drivers to be more important, we can lower the weight of the correlation score.

```{r Creating a composite score }

corr_weight <- 0.5
score_nSales <- merge(corr_nSales, magnitude_nSales, by = 
                        c("Store1", "Store2"))[, scoreNSales := (corr_measure + mag_measure)/2]
score_nSales[order(-scoreNSales)]


score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = 
                            c("Store1", "Store2"))[, scoreNCust := (corr_measure + mag_measure)/2]
score_nCustomers[order(-scoreNCust)]
```

Now we have a score for each of total number of sales and number of customers. 
Let's combine the two via a simple average.

```{r}
score_Control <- merge(score_nSales, score_nCustomers, by = 
                         c("Store1", "Store2"))

score_Control[, finalControlScore := scoreNSales * 0.5 + scoreNCust * 0.5]

score_Control[order(-finalControlScore)]


```

The store with the highest score is then selected as the control store since it is most similar to the trial store. From our results we can see that 233 can be selected as control store. Lets confirm that

```{r Selecting Control Store}
control_store <- score_Control[Store1 == trial_store, ][order(-finalControlScore)][2, Store2]
control_store
```


## Including Plots

Now that we have found a control store, let's check visually if the drivers are indeed similar in the period before the trial.
We'll look at total sales first.

```{r pressure, echo=FALSE}
#### Visual checks on trends based on the drivers
measureOverTimeSales <- measureOverTime
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
 ifelse(STORE_NBR == control_store, "Control", "Other stores")) ][, totSales := mean(totSales), by = c("YEARMONTH", 
"Store_type")][, TransactionMonth := as.Date(paste(YEARMONTH %/% 
100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")][YEARMONTH < 201903 , ]

```

```{r}
pastSales
```

```{r Plot 1}
ggplot(pastSales, aes(TransactionMonth, totSales, color = Store_type)) +
 geom_line() +labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```


Next, number of customers.
```{r pressure2, echo=FALSE}
#### Visual checks on trends based on the drivers
measureOverTimeCusts <- measureOverTime
pastCusts <- measureOverTimeCusts[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
 ifelse(STORE_NBR == control_store, "Control", "Other stores")) ][, numberCustomers := mean(nCustomers), by = c("YEARMONTH", 
"Store_type")][, TransactionMonth := as.Date(paste(YEARMONTH %/% 
100, YEARMONTH %% 100, 1, sep = "-"), "%Y-%m-%d")][YEARMONTH < 201903 , ]

```

```{r}
unique(pastCusts)
```

```{r Plot 2}
ggplot(pastCusts, aes(TransactionMonth, numberCustomers, color = Store_type)) +
 geom_line() +labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

The trial period goes from the start of February 2019 to April 2019. We now want to see if there has been an uplift in overall chip sales. 
We'll start with scaling the control store's sales to a level similar to control for any differences between the two stores outside of the trial period. 
```{r Comparison of results during trial}
#### Scale pre-trial control sales to match pre-trial trial store sales 
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR == trial_store & 
                                                   YEARMONTH < 201902,sum(totSales)]/preTrialMeasures[STORE_NBR == control_store &
                                                   YEARMONTH < 201902, sum(totSales)]

#### Apply the scaling factor
measureOverTimeSales <- measureOverTime
scaledControlSales <- measureOverTimeSales[STORE_NBR == control_store, ][ , controlSales := totSales * scalingFactorForControlSales]
```

```{r}
scaledControlSales

measureOverTime[STORE_NBR == trial_store]
```
Now that we have comparable sales figures for the control store, we can calculate the percentage difference between the scaled control sales and the trial store's sales during the trial period.

```{r Percentage Difference}
#### Calculate the percentage difference between scaled control sales and trial sales
percentageDiff <- merge(scaledControlSales[, c("YEARMONTH", "controlSales")],
                        measureOverTime[STORE_NBR == trial_store, c("totSales", "YEARMONTH")],
                        by = "YEARMONTH")[, percentageDiff := abs(controlSales - totSales)/controlSales]

```

```{r}
percentageDiff # between control store sales and trial store sales

```

Let's see if the difference is significant!

As our null hypothesis is that the trial period is the same as the pre-trial period, let's take the standard deviation based on the scaled percentage differencein the pre-trial period 
```{r Standard Dev}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
```


```{r}
#### Note that there are 8 months in the pre-trial period 
#### hence 8 - 1 = 7 degrees of freedom
degreesOfFreedom <- 7 
```

```{r}
percentageDiff[, tValue := (percentageDiff - 0)/stdDev
               ][, TransactionMonth := as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100, 1, 
                                                     sep = "‐"), "%Y‐%m‐%d")
][YEARMONTH < 201905 & YEARMONTH > 201901, .(TransactionMonth,tValue)]

```

```{r}
#### Find the 95th percentile of the t distribution with the appropriate
#### degrees of freedom to compare against
qt(0.95, df = degreesOfFreedom)
```

We can observe that the t-value is much larger than the 95th percentile value of the t-distribution for March and April - i.e. the increase in sales in the trial store in March and April is statistically greater than in the control store.

Let's create a more visual version of this by plotting the sales of the control store, the sales of the trial stores and the 95th percentile value of sales of the control store.

```{r fig.align = "Center"}
measureOverTimeSales <- measureOverTime
#### Trial and control store total sales
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial", ifelse(STORE_NBR == 
             control_store, "Control", "Other stores"))][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")
             ][, TransactionMonth := as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100, 1, sep = "‐"), "%Y‐%m‐%d")
             ][Store_type %in% c("Trial", "Control"), ]

#### Control store 95th percentile
pastSales_Controls95 <- pastSales[Store_type == "Control", ][, totSales := totSales * (1 + stdDev * 2)
                        ][, Store_type := "Control 95th % confidence interval"]

#### Control store 5th percentile
pastSales_Controls5 <- pastSales[Store_type == "Control", ][, totSales := totSales * (1 - stdDev * 2)
                       ][, Store_type := "Control 5th % confidence interval"]

trialAssessment <- rbind(pastSales, pastSales_Controls95, pastSales_Controls5)

```

```{r Plot 3}
ggplot(trialAssessment, aes(TransactionMonth, totSales, color = Store_type)) + 
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,], aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 , ymax = Inf, color = NULL), show.legend = FALSE) +
 geom_line() +
 labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

The results show that the trial in store 77 is significantly different to its control store in the trial period as the trial store performance lies outside the 5% to 95% confidence interval of the control store in two of the three trial 
months.
Let's have a look at assessing this for number of customers as well.

```{r}
#### Scale pre-trial control sales to match pre-trial trial store sales 
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR == trial_store & YEARMONTH < 201902, sum(nCustomers)] /
  preTrialMeasures[STORE_NBR == control_store & YEARMONTH < 201902, sum(nCustomers)]

#### Apply the scaling factor
measureOverTimeCusts <- measureOverTime

scaledControlCustomers <- measureOverTimeCusts[STORE_NBR == control_store,
][ , controlCustomers := nCustomers * scalingFactorForControlCust
][, Store_type := ifelse(STORE_NBR ==trial_store, "Trial",
ifelse(STORE_NBR == control_store,"Control", "Other stores"))]
```

```{r}
percentageDiff <- merge(scaledControlCustomers[, c("YEARMONTH", "controlCustomers")],
measureOverTimeCusts[STORE_NBR == trial_store,c("nCustomers", "YEARMONTH")],
by = "YEARMONTH"
)[, percentageDiff := abs(controlCustomers-nCustomers)/controlCustomers]
```


Let's again see if the difference is significant visually!
```{r , fig.align = "Center"}
#### As our null hypothesis is that the trial period is the same as the pre-trial period, let's take the standard deviation based on the scaled percentage difference in the pre-trial period 

stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff]) 
degreesOfFreedom <- 7 
#### Trial and control store number of customers
pastCustomers <- measureOverTimeCusts[, nCusts := mean(nCustomers), by =
c("YEARMONTH", "Store_type")
 ][Store_type %in% c("Trial", "Control"), ]
#### Control store 95th percentile
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control",
 ][, nCusts := nCusts * (1 + stdDev * 2)
 ][, Store_type := "Control 95th % confidence interval"]
#### Control store 5th percentile
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control",
 ][, nCusts := nCusts * (1 - stdDev * 2)
 ][, Store_type := "Control 5th % confidence interval"]

trialAssessment <- rbind(pastCustomers, pastCustomers_Controls95,
pastCustomers_Controls5)

```

```{r Plot 4}
ggplot(trialAssessment, aes(TransactionMonth, nCusts, color = Store_type)) +
 geom_rect(data = trialAssessment, aes(xmin = min(TransactionMonth) , xmax = max(TransactionMonth) , ymin = 0, ymax = Inf, 
 color = NULL), show.legend = FALSE) +
 geom_line() +
 labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month") 

```


Great, they look visuslly similar.Now let's repeat finding the control store and assessing the impact of the trial for each of the other two trial stores. 


## Trial Store 86


```{r Measure Over Time}
measureOverTime86 <- Data[, .(totSales = sum(TOT_SALES),
                            nCustomers = uniqueN(LYLTY_CARD_NBR),
                            nTxnPerCust = (uniqueN(TXN_ID))/(uniqueN(LYLTY_CARD_NBR)),
                            nChipsPerTxn = (sum(PROD_QTY))/(uniqueN(TXN_ID)) , 
                            avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY) ) , by = c("STORE_NBR", "YEARMONTH")][order(STORE_NBR, YEARMONTH)]

measureOverTime86

```


Calculating magnitude and correlation for sales and customers 

```{r Corr and Magn}
trial_store <- 86

corr_nSales <- calculateCorrelation(preTrialMeasures, quote(totSales),trial_store)
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales), trial_store)

corr_nCustomers <- calculateCorrelation(preTrialMeasures, quote(nCustomers), trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures, quote(nCustomers), trial_store)
```

```{r}
corr_nSales[order(-corr_measure)]

corr_nCustomers[order(-corr_measure)]
```

```{r}
magnitude_nSales[order(-mag_measure)]

magnitude_nCustomers[order(-mag_measure)]
```


```{r Creating a composite score 2}
score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))[ , scoreNSales := (corr_measure + mag_measure)/2]
score_nSales[order(-scoreNSales)]

score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))[ , scoreNCust := (corr_measure + mag_measure)/2]
score_nCustomers[order(-scoreNCust)]
```


```{r Selecting Control Store 2}
score_Control <- merge(score_nSales, score_nCustomers, by = c("Store1","Store2"))
score_Control[, finalControlScore := scoreNSales * 0.5 + scoreNCust * 0.5]
#### Select control stores based on the highest matching store
#### (closest to 1 but not the store itself, i.e. the second ranked highest store)
#### Select control store for trial store 86
control_store <- score_Control[Store1 == trial_store, ][order(-finalControlScore)][2, Store2]
control_store

```
Looks like store 155 will be a control store for trial store 86. Again, let's check visually if the drivers are indeed similar in the period before the trial.
We'll look at total sales first.


### Including Plots

Now that we have found a control store, let's check visually if the drivers are indeed similar in the period before the trial.
We'll look at total sales first.

```{r Sales Measures}
measureOverTimeSales <- measureOverTime86
pastSales <- measureOverTimeSales[, Store_type:= ifelse(STORE_NBR == trial_store, "Trial", ifelse(STORE_NBR== control_store, "Control", "Other stores"))][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")][, TransactionMonth:= as.Date(paste(YEARMONTH%/%100, YEARMONTH%% 100, 1, sep = "-"), "%Y-%m-%d")][YEARMONTH <210903] 
```



```{r Plot 5}
ggplot(pastSales, aes(TransactionMonth, totSales, color = Store_type)) +
geom_line(aes(linetype = Store_type)) +
labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

Great, sales are trending in a similar way.
Next, number of customers.

```{r Customer Measures}
measureOverTimeCusts <- measureOverTime86
pastCustomers <- measureOverTimeCusts[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, numberCustomers := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%
                                        100, YEARMONTH %% 100, 1, sep = "‐"), "%Y‐%m‐%d")
][YEARMONTH < 201903 , ]

ggplot(pastCustomers, aes(TransactionMonth, numberCustomers, color = Store_type)) + 
  geom_line() + labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

Good, the trend in number of customers is also similar. Let's now assess the impact of the trial on sales.

```{r Scale control}
#### Scale pre‐trial control sales to match pre‐trial trial store sales
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR ==
control_store & YEARMONTH < 201902, sum(totSales)]
#### Apply the scaling factor
measureOverTimeSales <- measureOverTime86
scaledControlSales <- measureOverTimeSales[STORE_NBR == control_store, ][ ,controlSales := totSales * scalingFactorForControlSales]
#### Calculate the percentage difference between scaled control sales and trial sales
percentageDiff <- merge(scaledControlSales[, c("YEARMONTH", "controlSales")],
measureOverTime[STORE_NBR == trial_store, c("totSales", "YEARMONTH")],
by = "YEARMONTH"
)[, percentageDiff := abs(controlSales-totSales)/controlSales]

```

As our null hypothesis is that the trial period is the same as the pre-trial period, let's take the standard deviation based on the scaled percentage difference in the pre-trial period.

```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7

measureOverTimeSales <- measureOverTime
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%100, YEARMONTH %% 100, 1, sep = "‐"), "%Y‐%m‐%d")
][Store_type %in% c("Trial", "Control"), ]

#### Control store 95th percentile
pastSales_Controls95 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 + stdDev * 2)
][, Store_type := "Control 95th % confidence interval"]

#### Control store 5th percentile
pastSales_Controls5 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 - stdDev * 2)
][, Store_type := "Control 5th % confidence interval"]
trialAssessment <- rbind(pastSales, pastSales_Controls95, pastSales_Controls5)

```

```{r Plot 6}
ggplot(trialAssessment, aes(TransactionMonth, totSales, color = Store_type)) +
geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 ,
ymax = Inf, color = NULL), show.legend = FALSE) + 
  geom_line(aes(linetype = Store_type)) + 
  labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```

The results show that the trial in store 86 is not significantly different to its control store in the trial period as the trial store performance lies inside the 5% to 95% confidence interval of the control store in two of the three trial months.
Let's have a look at assessing this for the number of customers as well.


```{r}
#### Scale pre-trial control customers to match pre-trial trial store customers
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(nCustomers)]/preTrialMeasures[STORE_NBR ==
control_store & YEARMONTH < 201902, sum(nCustomers)]
#### Apply the scaling factor
measureOverTimeCusts <- measureOverTime86
scaledControlCustomers <- measureOverTimeCusts[STORE_NBR == control_store,
][ , controlCustomers := nCustomers * scalingFactorForControlCust
][, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store,"Control", "Other stores"))
]
```

```{r Calculate Percentage Diff}
#### Calculate the percentage difference between scaled control sales and trial sales

percentageDiff <- merge(scaledControlCustomers[, c("YEARMONTH","controlCustomers")],measureOverTime[STORE_NBR == trial_store, c("nCustomers", "YEARMONTH")],
by = "YEARMONTH")[, percentageDiff := abs(controlCustomers-nCustomers)/controlCustomers]
```

```{r}
percentageDiff
```

```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7
#### Trial and control store number of customers
pastCustomers <- measureOverTimeCusts[, nCusts := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][Store_type %in% c("Trial", "Control"), ]
#### Control store 95th percentile
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control",
][, nCusts := nCusts * (1 + stdDev * 2)
][, Store_type := "Control 95th % confidence interval"]
#### Control store 5th percentile
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control",
][, nCusts := nCusts * (1 - stdDev * 2)
][, Store_type := "Control 5th % confidence interval"]
trialAssessment <- rbind(pastCustomers, pastCustomers_Controls95,pastCustomers_Controls5)
```

```{r Plot 7}
ggplot(trialAssessment, aes(TransactionMonth, nCusts, color = Store_type)) +
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 ,
ymax = Inf, color = NULL), show.legend = FALSE) + geom_line() + 
labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```


It looks like the number of customers is significantly higher in all of the three months. This seems to suggest that the trial had a significant impact on increasing the number of customers in trial store 86 but as we saw, sales were not significantly higher. We should check with the Category Manager if there were special deals in the trial store that were may have resulted in lower prices, 
impacting the results.


## Trial Store 88

Calculating measures over time for store 88

```{r Measure Over Time 2}
measureOverTime88 <- Data[, .(totSales = sum(TOT_SALES),
nCustomers = uniqueN(LYLTY_CARD_NBR),
nTxnPerCust = uniqueN(TXN_ID)/uniqueN(LYLTY_CARD_NBR),
nChipsPerTxn = sum(PROD_QTY)/uniqueN(TXN_ID),
avgPricePerUnit = sum(TOT_SALES)/sum(PROD_QTY))
, by = c("STORE_NBR", "YEARMONTH")][order(STORE_NBR, YEARMONTH)]

```


```{r Corr and Magn 2}
trial_store <- 88

corr_nSales <- calculateCorrelation(preTrialMeasures, quote(totSales),trial_store)
magnitude_nSales <- calculateMagnitudeDistance(preTrialMeasures, quote(totSales), trial_store)

corr_nCustomers <- calculateCorrelation(preTrialMeasures, quote(nCustomers), trial_store)
magnitude_nCustomers <- calculateMagnitudeDistance(preTrialMeasures, quote(nCustomers), trial_store)
```

```{r}
corr_nSales[order(-corr_measure)]

corr_nCustomers[order(-corr_measure)]
```

```{r}
magnitude_nSales[order(-mag_measure)]

magnitude_nCustomers[order(-mag_measure)]
```


Create a combined score composed of correlation and magnitude by merging the correlations table and the magnitudes table, for each driver.

```{r Creating a composite score 3}
score_nSales <- merge(corr_nSales, magnitude_nSales, by = c("Store1", "Store2"))[ , scoreNSales := (corr_measure + mag_measure)/2]

score_nCustomers <- merge(corr_nCustomers, magnitude_nCustomers, by = c("Store1", "Store2"))[ , scoreNCust := (corr_measure + mag_measure)/2]

```


Select control stores based on the highest matching store (closest to 1 but not the store itself, i.e. the second ranked highest store)

```{r Selecting Control Store 3}
score_Control <- merge(score_nSales, score_nCustomers, by = c("Store1","Store2"))
score_Control[, finalControlScore := scoreNSales * 0.5 + scoreNCust * 0.5]

control_store <- score_Control[Store1 == trial_store, ][order(-finalControlScore)][2, Store2]
control_store

```

We've now found store 237 to be a suitable control store for trial store 88. Again, let's check visually if the drivers are indeed similar in the period before the trial.

### Including Plots

Now that we have found a control store, let's check visually if the drivers are indeed similar in the period before the trial.
We'll look at total sales first.

```{r Sales Measures 2}
measureOverTimeSales <- measureOverTime88
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, totSales := mean(totSales), by = c("YEARMONTH","Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/% 100, YEARMONTH %% 100, 1, sep = "‐"), "%Y‐%m‐%d")
][YEARMONTH < 201903 , ]

#### Plotting this in a graph
ggplot(pastSales, aes(TransactionMonth, totSales, color = Store_type)) +
geom_line(aes(linetype = Store_type)) +
labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")

```


Great, the trial and control stores have similar total sales.
Next, number of customers.

```{r Customer Measures 2}
measureOverTimeCusts <- measureOverTime88
pastCustomers <- measureOverTimeCusts[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, numberCustomers := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%
                                        100, YEARMONTH %% 100, 1, sep = "‐"), "%Y‐%m‐%d")
][YEARMONTH < 201903 , ]

#### Plotting this in a graph
ggplot(pastCustomers, aes(TransactionMonth, numberCustomers, color = Store_type)) + 
  geom_line() + labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")

```


Total number of customers of the control and trial stores are also similar.
Let's now assess the impact of the trial on sales.


```{r Scale Control 2}
#### Scale pre‐trial control sales to match pre‐trial trial store sales
scalingFactorForControlSales <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(totSales)]/preTrialMeasures[STORE_NBR ==
control_store & YEARMONTH < 201902, sum(totSales)]

#### Apply the scaling factor
measureOverTimeSales <- measureOverTime88
scaledControlSales <- measureOverTimeSales[STORE_NBR == control_store, ][ ,controlSales := totSales * scalingFactorForControlSales]
```

```{r Calculate Percentage Diff 2}
#### Calculate the percentage difference between scaled control sales and trial sales
percentageDiff <- merge(scaledControlSales[, c("YEARMONTH", "controlSales")],measureOverTime[STORE_NBR == trial_store, c("totSales", "YEARMONTH")],by = "YEARMONTH")[, percentageDiff := abs(controlSales-totSales)/controlSales]

```

```{r}
percentageDiff
```

```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7

measureOverTimeSales <- measureOverTime88
pastSales <- measureOverTimeSales[, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store, "Control", "Other stores"))
][, totSales := mean(totSales), by = c("YEARMONTH", "Store_type")
][, TransactionMonth := as.Date(paste(YEARMONTH %/%100, YEARMONTH %% 100, 1, sep = "‐"), "%Y‐%m‐%d")
][Store_type %in% c("Trial", "Control"), ]

#### Control store 95th percentile
pastSales_Controls95 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 + stdDev * 2)
][, Store_type := "Control 95th % confidence interval"]

#### Control store 5th percentile
pastSales_Controls5 <- pastSales[Store_type == "Control",
][, totSales := totSales * (1 - stdDev * 2)
][, Store_type := "Control 5th % confidence interval"]
trialAssessment <- rbind(pastSales, pastSales_Controls95, pastSales_Controls5)
```

```{r Plot 8}
#### Plotting these in one nice graph
ggplot(trialAssessment, aes(TransactionMonth, totSales, color = Store_type)) +
geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 ,
ymax = Inf, color = NULL), show.legend = FALSE) + 
  geom_line(aes(linetype = Store_type)) + 
  labs(x = "Month of operation", y = "Total sales", title = "Total sales by month")
```


The results show that the trial in store 88 is significantly different to its control store in the trial period as the trial store performance lies outside of the 5% to 95% confidence interval of the control store in two of the three trial 
months.
Let's have a look at assessing this for number of customers as well.

```{r}
scalingFactorForControlCust <- preTrialMeasures[STORE_NBR == trial_store &
YEARMONTH < 201902, sum(nCustomers)]/preTrialMeasures[STORE_NBR ==
control_store & YEARMONTH < 201902, sum(nCustomers)]
#### Apply the scaling factor
measureOverTimeCusts <- measureOverTime88
scaledControlCustomers <- measureOverTimeCusts[STORE_NBR == control_store,
][ , controlCustomers := nCustomers * scalingFactorForControlCust
][, Store_type := ifelse(STORE_NBR == trial_store, "Trial",
ifelse(STORE_NBR == control_store,"Control", "Other stores"))
]

```


```{r}
percentageDiff <- merge(scaledControlCustomers[, c("YEARMONTH","controlCustomers")],measureOverTime[STORE_NBR == trial_store, c("nCustomers", "YEARMONTH")],
by = "YEARMONTH")[, percentageDiff := abs(controlCustomers-nCustomers)/controlCustomers]
```

```{r}
percentageDiff
```



```{r}
stdDev <- sd(percentageDiff[YEARMONTH < 201902 , percentageDiff])
degreesOfFreedom <- 7

#### Trial and control store number of customers
pastCustomers <- measureOverTimeCusts[, nCusts := mean(nCustomers), by = c("YEARMONTH", "Store_type")
][Store_type %in% c("Trial", "Control"), ]

#### Control store 95th percentile
pastCustomers_Controls95 <- pastCustomers[Store_type == "Control",
][, nCusts := nCusts * (1 + stdDev * 2)
][, Store_type := "Control 95th % confidence interval"]

#### Control store 5th percentile
pastCustomers_Controls5 <- pastCustomers[Store_type == "Control",
][, nCusts := nCusts * (1 - stdDev * 2)
][, Store_type := "Control 5th % confidence interval"]
trialAssessment <- rbind(pastCustomers, pastCustomers_Controls95,pastCustomers_Controls5)

#### Plotting these in one nice graph
ggplot(trialAssessment, aes(TransactionMonth, nCusts, color = Store_type)) +
  geom_rect(data = trialAssessment[ YEARMONTH < 201905 & YEARMONTH > 201901 ,],
aes(xmin = min(TransactionMonth), xmax = max(TransactionMonth), ymin = 0 ,
ymax = Inf, color = NULL), show.legend = FALSE) + geom_line() + 
labs(x = "Month of operation", y = "Total number of customers", title = "Total number of customers by month")
```

Total number of customers in the trial period for the trial store is significantly higher than the control store for two out of three months, which indicates a positive trial effect.


## CONCLUSION

We've found control stores 233, 155, 237 for trial stores 77, 86 and 88 respectively. The results for trial stores 77 and 88 during the trial period show a significant difference in at least two of the three trial months but this is not the case for trial store 86. We can check with the client if the implementation of the trial was different in trial store 86 but overall, the trial shows a significant increase in sales. Now that we have finished our analysis, we can prepare our presentation to the Category Manager.



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
